### 2022-06-28 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@abd9f48a0e70ba1f1619e6a9892a9eeb1e1aefc9 🚀

Overall, these results suggest that the boundaries of bat and
betacoronavirus biogeographic regions are largely consistent. This may
be surprising, given that cross-species transmission may play a stronger
role in coronavirus diversification than
cospeciation---\textsuperscript{2}a property that would theoretically
allow for substantial broad divergence in their biogeography. However,
host jumps at the family level or higher are relatively rare and
significant events in coronavirus evolutionary
history;\textsuperscript{2,13} as a result, the mosaic of
betacoronavirus phylogeography is assembled from a set of overlapping
smaller coevolutionary systems, superimposed in space and filtered by
the importance of different subgroups in local host communities. For
example, the most speciose and cosmopolitan family of bats, the vesper
bats (Vespertilionidae), are considered the primary hosts of the
subgenus \emph{Merbecovirus} (MERS-like viruses);\textsuperscript{3,13}
but in the Americas, where merbecoviruses are the only lineage present,
they have only been found in other bat taxa (e.g., Molossidae,
Phyllostomidae) (refs1). At the coarsest scale, these heterogeneities
are lost, and betacoronavirus biogeography tracks the deep rifts in bat
evolutionary history---but within broad regions, the component
coevolutionary systems may have very different dynamics.

@@

Bats, the second most diverse groups of mammals, are found worldwide;
gradients in their species richness generally track broader patterns of
mammal diversity, with a striking Neotropical hotspot (especially in the
Amazon basin) and a secondary hotspot centered in the southeast Asian
peninsula. These hotspots of bat diversity are generally presumed to be
hotspots of viral adaptive radiation, and therefore areas of concern for
human health.\textsuperscript{2,14} However, the hotspots of known bat
betacoronavirus hosts show a distinct pattern, with primary hotspots
(both in terms of size and higher values) of host richness situated in
southeast Asia, parts of southern Europe, and to a lesser extent parts
of Africa in the -25-0 range of latitudes (fig.~\ref{fig:richness};
top). Although hundreds of species likely host undiscovered
betacoronaviruses, machine learning predictions have suggested that
these undiscovered reservoirs should follow the same diversity
gradient.\textsuperscript{15} In principle, these hotspots of
locally-diverse, virus-rich bat communities should drive more adaptive
diversification in their viruses.

@@

However, we find that the global pattern of betacoronavirus phylogenetic
distinctiveness is quite distinct from both bat host richness and
phylogenetic distinctiveness (fig.~\ref{fig:richness}; bottom). In
contrast to the sparsity of Neotropical betacoronavirus hosts, South and
Central America have the most evolutionary distinct hosts \emph{and}
viruses, followed by secondary hotspots in southeast Asia and the Rift
Valley region have mostly distinct viruses. Some degree of sampling bias
may contribute to these patterns: for example, the Neotropics are one of
the places where the fewest bat betacoronavirus sequences have been
generated (cite2), resulting in a sparser phylogenetic tree, and
artificially inflating distinctiveness; conversely, disproportionate
research effort in eastern China\textsuperscript{16} may have led to a
more complete inventory of the local diversity of coronaviruses, again
inflating these metrics relative to underlying patterns. Even accounting
for these potential biases, though, there is obvious heterogeneity in
betacoronavirus evolutionary distinctiveness that is distinct from
overall bat diversity.

Overall, these patterns recapitulate the evolutionary history of both
the order Chiroptera and the genus \emph{Betacoronavirus}. Horseshoe
bats (Rhinolophidae) include the reservoirs of the SARS-like viruses
(subgenus \emph{Sarbecovirus}), the group of pandemic threats that have
been of the greatest interest to researchers\textsuperscript{13} (and so
have been sampled most intensively).\textsuperscript{16} The hotspots of
host richness and viral diversity in southeast Asia---both of which are
disproportionately high, considering the global landscape of bat species
richness---are almost entirely driven by viral adaptive radiation
through host switching within this clade\textsuperscript{3,15}. In
contrast, the Neotropical hotspot of viral distinctiveness is driven by
isolation by host vicariance. Out of the four main groups of
betacoronaviruses, only merbecoviruses have been found in animals in the
Americas--- an introduction that is generally presumed to be
ancient.\textsuperscript{3,17} While comparatively understudied, New
World merbecoviruses have been found in the ghost-faced bats
(Mormoopidae), Neotropical leaf-nosed bats (Phyllostomidae), and
free-tailed bats (Molossidae) (refs3). The former two groups and a clade
of the latter are endemic to the Neotropics, while the explosive
adaptive radiations of the phyllostomids are responsible for the hotspot
of bat diversity in the Amazon (refs4). Together, these clades of New
World bats play host to a distinct regime of betacoronavirus
coevolution.

@@

As host richness, joint distinctiveness, or phylogeographic structure
suggest that the bat-betacoronaviruses complex is globally fragmented
enough to give rise to both different levels of risk (as evidenced by
the spatial location of spillover events) and different types of
co-evolutionary dynamics, we turn to the Geographic Mosaic Theory of
Coevolution to provide a measure of risk accounting for multiple
processes. In fig.~\ref{fig:trivariate}, we overlapped three components
of spillover risk: viral sharing, \emph{i.e.} the chance that two bats
will share viruses overall; Local Contribution to Beta Diversity,
\emph{i.e.} the fact that a bat community is compositionally unique
compared to the average compositional similarity across the entire
system; finally, host phylogenetic diversity, \emph{i.e.} how dispersed
the bats in a location are within the tree of life. This approach leads
to the definition of broad biogeographic regions of risk, where the same
color represents the same type of risk. By way of constrat to figures
fig.~\ref{fig:richness} and fig.~\ref{fig:biogeo}, these regions do not
necessarilly overlap with previous spatial partitions of the
bat-betacoronaviruses complex.

\begin{figure}
\hypertarget{fig:trivariate}{%
\centering
\includegraphics{figures/risk_trivariate.png}
\caption{Trivariate additive mapping of the components of risk in the
red/green/blue, where high virus sharing is encoded in the blue channel,
host phylogenetic diversity in the green channel, and compositional
uniqueness in the red channel. A pixel that would maximize all measures
(highest possible risk) would be a pure white (specifically RGB(1.0,
1.0. 1.0)), and a pixel with the lowest possible values would be pure
black (specifically RGB(0.0, 0.0, 0.0)). Therefore, lighter values (the
sum of the three channels gets closer to 3) indicate higher risk, and
the color indicates the proportional distribution of the factors making
up the total risk.}\label{fig:trivariate}
}
\end{figure}

From the perspective of spillover risk, the most important combination
of factors is a high phylogenetic diversity of hosts with low viral
sharing; this, essentially, means that very different betacoronaviruses
could co-exist within the same place. This is particularly the case
given that betacoronaviruses often evolve and even achieve host shifts
through recombination, which requires the co-occurrence of sufficiently
distinct viruses to be a major driver of emergence. In
fig.~\ref{fig:trivariate}, this corresponds to yellow to pale green
areas, which are essentially limited to South-Eastern Asia, and to some
part of Sub-Saharan Africa. Adopting a geographic mosaic theory
perspective on risk, other regions of the world are of lesser concern
(fig.~\ref{fig:risk}). Our risk decomposition does not account for viral
diversity or distinctiveness. The simple rationale behind it is that the
acquisition of viral data is rarely disconnected from the acquisition of
host data. There are more sources of information on hosts than on
viruses, allowing to develop a host-centric perspective on risk
(although this estimate would more accurate with viral traits related to
\emph{e.g.} ability to switch hosts or pathogenic potential). Areas with
high bat diversity and high turnover \emph{may} facilitate the
evolutionary radiation of viruses, matching previous findings that the
diversification of bat coronaviruses is driven largely by host shifts
(inter-genus or higher levels of cross-species transmission) and, to a
lesser degree, cospeciation and sharing, representing intra-genus
cross-species transmission.\textsuperscript{2} This diversification is
not an actual risk factor for spillover itself, but acts downstream of a
spillover event by increasing the random chance of the emergence of a
virus with the raw genomic components required for the potential to
infect humans.

\begin{figure}
\hypertarget{fig:risk}{%
\centering
\includegraphics{figures/risk_map.png}
\caption{Extraction of a measure of \emph{Betacoronavirus} spillover
risk from bat hosts based on the colorimetric space from
fig.~\ref{fig:trivariate}. The risk is a composite measure of the color
value and angular distance to the yellow hue, as defined in the methods,
ranged in the unit space. Based on this analyses, regions at high risk
of spillover are southeast Asia and Madagascar.}\label{fig:risk}
}
\end{figure}

From another perspective, areas of high host uniqueness and virus
sharing (red-to-pink) could provide hotspots of \emph{Betacoronavirus}
risk through mixing of unique viruses (via codivergence) and in turn
recombination. Under our framework, such a hotspot was identified in
Madagascar, where most bat species are endemic following evolutionary
divergence from sister species in both African and Asian
continents.\textsuperscript{18} Recent surveillance\textsuperscript{19}
has identified a novel \emph{Betacoronavirus} (in the subgenus
\emph{Nobecovirus}) in Madagascar-endemic pteropid bat species
(\emph{Pteropus rufus}, \emph{Rousettus madagascariensis}), emphasizing
strong proof of principle in model predictions.

@@

The relationship between the underlying pathogen pool and emergence risk
is mediated by both human-wildlife interfaces (the probability of
spillover) and opportunities for onward transmission (the probability
that spillovers become epidemics)\textsuperscript{1}. As a proxy for
both, we finally overlaid the risk component from the composite map (see
above) with the proportion of built land, as a proxy for a mix of
habitat disturbance, potential for bat synanthropy or contact with
bridge hosts like livestock,\textsuperscript{20,21} and human population
density and connectivity\textsuperscript{1,22,23}
(fig.~\ref{fig:compound}). Accounting for these factors, most of South
America and Europe are at comparatively lower risk, as--although densely
populated--settlements tend to be in areas with lower potential risk.
Conversely, regions like Malaysia and the northern coast of Australia
have a high evolutionary risk component, but should represent a
relatively lower effective risk due to low human density. However,
southeast Asia, the Indian subcontinent, and scattered hotspots in
sub-Saharan Africa are at high risk due to the overlap between human
populations and natural opportunities for cross-species transmission of
betacoronaviruses.

@@

Reassuringly, these predictions correspond to the geographic origins of
the three bat-origin coronaviruses that have recently emerged in human
populations. While available information puts the spillover of
SARS-CoV-2 in a live animal market in Wuhan, China, the ultimate origin
of the virus is almost certainly in a divergent lineage of
sarbecoviruses from the Indochinese peninsula that was poorly
characterized prior to the pandemic.\textsuperscript{24--26} Similarly,
the SARS-CoV outbreak began in Guangdong province in 2002, reaching
humans through small carnivore bridge hosts, but was eventually traced
back to a set of likely progenitor viruses found in cave-dwelling
horseshoe bats in Yunnan province;\textsuperscript{27} nearby, antibody
evidence has indicated human exposure to SARS-like
viruses.\textsuperscript{28} MERS-CoV was originally detected in Saudi
Arabia, accompanied by a nearly identical virus sequenced from an
Egyptian tomb bat (\emph{Taphozous perforatus})\textsuperscript{29}, but
is widespread in camels in East Africa and the Middle East, and may have
reached its bridge host decades earlier than originally
supposed;\textsuperscript{30} as a result, the geography of the original
bat-to-camel transmission is still widely regarded as uncertain. All of
these are broadly consistent with the risk factors we identify. Notably,
India and west Africa are additional hotspots that have yet to
experience the emergence of a bat coronavirus into human populations,
but may still be at risk---particularly given known gaps in bat
surveillance,\textsuperscript{16} and a dense population in both regions
with global connectivity. In any of these regions, surveillance on viral
reservoirs can be paired with targeted monitoring of high-risk human
populations (i.e., those with regular wildlife
contact)\textsuperscript{31} for maximum impact.

@@

Bats are important reservoir hosts for different classes of
microorganisms, many of which a threat to human
health.\textsuperscript{32,33} Chiropterans emerged around 64 million
years ago and are one of the most diverse mammalian orders, with an
estimated richness of more than 1400 species.\textsuperscript{34,35}
They exhibit a broad variety of habitat use, behaviour, and feeding
strategies, putting them at key positions in the delivery and
provisioning of several ecosystem services, tied to important
ecosystem-derived benefits to human.\textsuperscript{36} For example,
bats are an essential component of many seed-dispersal
networks.\textsuperscript{37} Over two-thirds of bats are know to be
either obligate or facultative insectivores, therefore actively
contributing for agricultural pest control,\textsuperscript{38,39} and
vectors of pathogens that put a risk on human
health.\textsuperscript{40,41} Because bats are globally distributed and
have a long evolutionary history, phylogeographic and biogeographic
approaches are required to shed light on the contemporary distribution
of coevolutionary processes between bats and the pathogens they host.
Not all areas in which bats, viruses, and human are co-occuring are
facing a risk of spillover towards human populations, and the areas in
which this risk exist may not be facing risks of the same nature and
magnitude.

Here, we propose a simple freamework with broad explanatory power that
helps contextualize discoveries like highly divergent nobecoviruses in
Madagascar and the previously-neglected adaptive radiation of
sarbecoviruses outside of southern China and throughout southeast Asia.
In doing so, it advances ecological theory beyond the current state of
the art for global maps of emergence risk. For example, previous studies
that have used host richness as proxy have predicted a high diversity of
unsampled bat viruses,\textsuperscript{14} bat
coronaviruses,\textsuperscript{2} and even specifically
betacoronaviruses\textsuperscript{15} in both the Amazon and southeast
Asia. While we find that both regions are characterized by highly
divergent host and viral communities, our framework identifies key
differences between the regions. We find that Latin America is a hotspot
of both host and viral distinctiveness, suggesting that this branch of
the bat-betacoronavirus complex may be undergoing independent
evolutionary dynamics from the rest of the global pool, but with limited
potential for viral diversification--- a finding that is supported by
previous work indicating a higher rate of codivergence in Latin
America.\textsuperscript{2} In contrast, in southeast Asia, host
richness and viral distinctiveness are high but sharing is low; this
suggests a different type of evolutionary dynamics that could generate
high local diversity of viruses through host switching and viral
recombination (see \emph{e.g.},\textsuperscript{13} as well as the
discovery of recombinant viruses that share genetic material from both
the SARS-CoV and SARS-CoV-2 branches of the Sarbecovirus
lineage).\textsuperscript{42} Both of these regions are priority areas
for sampling, especially given predictions that they contain many bat
hosts of undiscovered betacoronaviruses.\textsuperscript{15,16} However,
both the evolutionary and ecological aspects of emergence risk are
likely higher in southeast Asia---a fact that will only become more
relevant, as bats track shifting climates and exchange viruses with
other species, creating a hotspot of cross-species transmission unique
to the region.\textsuperscript{43}

The diversity and diversification potential of bats responds to
anthropogenic factors others than shifting climates.\textsuperscript{44}
Land use changes could significantly decrease bat suitability, notably
through effects on diet and availability of
habitats.\textsuperscript{45} As our results establish that the
diversification of bats betacoronaviruses happens on top of processes
affecting hosts, biogeographic variation in human population density and
anthropogenic disturbances may feed into co-evolutionary dynamics.
Increase in humans-hosts contacts also increase the risk of emergence of
novel diseases,\textsuperscript{46} so does the changes in landscape
connectivity at local/regional scales.\textsuperscript{47} This
represents a challenge for both conservation strategies and disease
ecology: some areas can a high emergence risk and more potential for the
acquisition of zoonotic viruses through bat-human
encounters.\textsuperscript{48} In particular, the challenge ahead lies
in the need to quantify actual exposure (and risk) accounting for
several transmission scenarios, including both direct and indirect bat -
human interactions, and feeding back into the provision of ecosystem
services by bats.

@@

We downloaded the data on bats hosts of \emph{Betacoronavirus} from
\texttt{https://www.viralemergence.org/betacov} on
Apr.~2022,\textsuperscript{15} and filtered it to ``known'' hosts
(established before the emergence of SARS-CoV-2) and ``novel'' hosts
(confirmed through sampling and competence assays since the initial data
collection). The original database was assembled by a combination of
data mining and literature surveys, including automated alerts on the
``bats'' and ``coronavirus'' keywords to identify novel empirical
evidence of bats-betacoronaviruses associations; this yielded a total of
126 known hosts, 47 of which were novel hosts.

@@

We downloaded the rangemap of every current bat species that was
classified as an empirically documented host of \emph{Betacoronavirus}
from the previous step, according to recent IUCN
data.\textsuperscript{49} The range maps were subsequently rasterized
using the \texttt{rasterize} function from
\texttt{GDAL}\textsuperscript{50} at a resolution of approximately
100kmx100km. For every pixel in the resulting raster where at least one
bat host of \emph{Betacoronavirus} was present, we extract the species
pool (list of all known bat hosts), which was used to calculate the
following risk assessment components: bat phylogenetic diversity, bat
compositional uniqueness, and predicted viral sharing risk.

@@

For every pixel, we measured Faith's Phylogenetic
Diversity\textsuperscript{51} based on a recent synthetic tree with
robust time calibration, covering about 6000 mammalian
species.\textsuperscript{52} Faith's PD measures the sum of unique
branches from an arbitrary root to a set of tips, and comparatively
larger values indicate a more phylogenetic diverse species pool. We
measured phylogenetic diversity starting from the root of the entire
tree (and not from Chiroptera); this bears no consequences on the
resulting values, since all branches leading up to Chiroptera are only
counted one per species pool, and (as we explain when describing the
assembly of the composite risk map), all individual risk components are
ranged in {[}0,1{]}. This measure incorporates a richness component,
which we chose not to correct for; the interpretation of the
phylogenetic diversity is therefore a weighted species richness, that
accounts for phylogenetic over/under-dispersal in some places.

>>>>

Overall, these results suggest that the boundaries of bat and
betacoronavirus biogeographic regions are largely consistent. This may
be surprising, given that cross-species transmission may play a stronger
role in coronavirus diversification than
cospeciation---\textsuperscript{2}a property that would theoretically
allow for substantial broad divergence in their biogeography. However,
host jumps at the family level or higher are relatively rare and
significant events in coronavirus evolutionary
history;\textsuperscript{2,13} as a result, the mosaic of
betacoronavirus phylogeography is assembled from a set of overlapping
smaller coevolutionary systems, superimposed in space and filtered by
the importance of different subgroups in local host communities. For
example, the most speciose and cosmopolitan family of bats, the vesper
bats (Vespertilionidae), are considered the primary hosts of the
subgenus \emph{Merbecovirus} (MERS-like viruses);\textsuperscript{3,13}
but in the Americas, where merbecoviruses are the only lineage present,
they have only been found in other bat taxa (\emph{e.g.}, Molossidae,
Phyllostomidae).\textsuperscript{14--17} At the coarsest scale, these
heterogeneities are lost, and betacoronavirus biogeography tracks the
deep rifts in bat evolutionary history---but within broad regions, the
component coevolutionary systems may have very different dynamics.

@@

Bats, the second most diverse groups of mammals, are found worldwide;
gradients in their species richness generally track broader patterns of
mammal diversity, with a striking Neotropical hotspot (especially in the
Amazon basin) and a secondary hotspot centered in the southeast Asian
peninsula. These hotspots of bat diversity are generally presumed to be
hotspots of viral adaptive radiation, and therefore areas of concern for
human health.\textsuperscript{2,18} However, the hotspots of known bat
betacoronavirus hosts show a distinct pattern, with primary hotspots
(both in terms of size and higher values) of host richness situated in
southeast Asia, parts of southern Europe, and to a lesser extent parts
of Africa in the -25-0 range of latitudes (fig.~\ref{fig:richness};
top). Although hundreds of species likely host undiscovered
betacoronaviruses, machine learning predictions have suggested that
these undiscovered reservoirs should follow the same diversity
gradient.\textsuperscript{19} In principle, these hotspots of
locally-diverse, virus-rich bat communities should drive more adaptive
diversification in their viruses.

@@

However, we find that the global pattern of betacoronavirus phylogenetic
distinctiveness is quite distinct from both bat host richness and
phylogenetic distinctiveness (fig.~\ref{fig:richness}; bottom). In
contrast to the sparsity of Neotropical betacoronavirus hosts, South and
Central America have the most evolutionary distinct hosts \emph{and}
viruses, followed by secondary hotspots in southeast Asia and the Rift
Valley region have mostly distinct viruses. Some degree of sampling bias
may contribute to these patterns: for example, the Neotropics are one of
the places where the fewest bat betacoronavirus sequences have been
generated (cite2), resulting in a sparser phylogenetic tree, and
artificially inflating distinctiveness; conversely, disproportionate
research effort in eastern China\textsuperscript{20} may have led to a
more complete inventory of the local diversity of coronaviruses, again
inflating these metrics relative to underlying patterns. Even accounting
for these potential biases, though, there is obvious heterogeneity in
betacoronavirus evolutionary distinctiveness that is distinct from
overall bat diversity.

Overall, these patterns recapitulate the evolutionary history of both
the order Chiroptera and the genus \emph{Betacoronavirus}. Horseshoe
bats (Rhinolophidae) include the reservoirs of the SARS-like viruses
(subgenus \emph{Sarbecovirus}), the group of pandemic threats that have
been of the greatest interest to researchers\textsuperscript{13} (and so
have been sampled most intensively).\textsuperscript{20} The hotspots of
host richness and viral diversity in southeast Asia---both of which are
disproportionately high, considering the global landscape of bat species
richness---are almost entirely driven by viral adaptive radiation
through host switching within this clade\textsuperscript{3,19}. In
contrast, the Neotropical hotspot of viral distinctiveness is driven by
isolation by host vicariance. Out of the four main groups of
betacoronaviruses, only merbecoviruses have been found in animals in the
Americas--- an introduction that is generally presumed to be
ancient.\textsuperscript{3,21} While comparatively understudied, New
World merbecoviruses have been found in the ghost-faced bats
(Mormoopidae), Neotropical leaf-nosed bats (Phyllostomidae), and
free-tailed bats (Molossidae).\textsuperscript{14--17} The former two
groups and a clade of the latter are endemic to the Neotropics, while
the explosive adaptive radiations of the phyllostomids are responsible
for the hotspot of bat diversity in the Amazon.\textsuperscript{22}
Together, these clades of New World bats play host to a distinct regime
of betacoronavirus coevolution.

@@

Both the diverse origins of betacoronavirus spillover events and the
existence of well defined co-phylogenetic regions suggest that the
system is spatially fragmented enough to create risks of emergence that
are different in their amplitude and in the coevolutionary processes
underlying them. Therefore, we turn to the Geographic Mosaic Theory of
Coevolution to provide an assesment of risk accounting for these
processes. In fig.~\ref{fig:trivariate}, we overlapped three components
of spillover risk: viral sharing (high viral sharing suggests slower,
diffuse coevolution); host phylogenetic diversity (high diversity
represents different evolutionary histories, suggesting more variation
in immune strategies); host community uniqueness (high uniqueness
suggests more potential for viruses to be exposed to novel host traits).
This approach leads to the definition of broad biogeographic regions of
risk, where close colors represent similar risks, and paler pixels
represent overall higher risk (see Methods). These regions do not neatly
overlap with those defined in fig.~\ref{fig:richness} or
fig.~\ref{fig:biogeo}, reinforcing the notion that local-scale
coevolutionary mosaics exist within co-phylogenetic regions.

\begin{figure}
\hypertarget{fig:trivariate}{%
\centering
\includegraphics{figures/risk_trivariate.png}
\caption{Trivariate additive mapping of the components of risk in the
red/green/blue, where high virus sharing is encoded in the blue channel,
host phylogenetic diversity in the green channel, and compositional
uniqueness in the red channel. The main driver of emergence risk
(possibility of spatially overlapped but coevolutionarily indepent
host-viral dynamics) corresponds to low viral sharing, \emph{i.e.}
pixels around yellow. Pixels in the yellow-green space (Africa)
correspond to areas where, despite the potential for viral
diversification, coevolutionary dynamics are likely to currently be
dominated by the effect of host phylogenetic
diversity.}\label{fig:trivariate}
}
\end{figure}

Emergence risk is maximized under low viral sharing (host-virus pairs
coevolve independantly), high phylogenetic diversity (viruses are
exposes to different host clades), and high host uniqueness (viruses are
experiencing novel, heterogeneous host traits combinations). Under these
conditions, very different betacoronaviruses could co-exist at the same
place, yet evolve in independent ways. As betacoronaviruses often evolve
(including host shifts) through recombination, the co-occurrence of
sufficiently distinct viruses is a sufficient major driver of emergence,
and the regions that meet these conditions therefore represent the
higher risk. In fig.~\ref{fig:trivariate}, this corresponds to yellow
areas (dynamics dominated by low viral sharing, with equal contributions
of selection mosaics and trait remixing; South-Eastern Asia, and the
Indian sub-continent), green-yellow areas (dynamics with low viral
sharing but dominated by the selection mosaic effect of host diversity;
Africa below the Sahara desert), and would correspond to red-yellow
areas (dynamics with low viral sharing but dominated by trait remixing
in host communities; Middle-East). Indeed, these regions are broadly
reflected in the risk map fig.~\ref{fig:risk}.

Under this framework, other regions (where high viral sharing dominates
the dynamics; see \emph{e.g.} Latin America, Eurasia above a northing of
30) represent a lower risk of emergence. Nevertheless, areas of high
host uniqueness coupled with high viral sharing(red-to-pink in
fig.~\ref{fig:trivariate}) could provide future hotspots of
\emph{Betacoronavirus} emregence risk through the suddent divergence of
currently diffuse coevolutionary dynamics; this is a likely scenario
knowing that can be facilitated by codivergence followed by
recombination. Madagascar, where most bat species are endemic following
evolutionary divergence from sister species in both African and Asian
continents,\textsuperscript{23} is one such potential future hotspot.
Indeed, recent surveillance\textsuperscript{24} has identified a novel
\emph{Betacoronavirus} (in the subgenus \emph{Nobecovirus}) in
Madagascar-endemic pteropid bat species (\emph{Pteropus rufus},
\emph{Rousettus madagascariensis}), emphasizing strong proof of
principle in model predictions.

Our risk decomposition does not account for viral diversity or
distinctiveness. Viral data acquisition is rarely disconnected from the
acquisition of host data. There are more sources of information on hosts
than on viruses, allowing to develop a more robust host-centric
perspective on risk. Any emergence risk estimate would benefit from
viral traits related to \emph{e.g.} ability to switch hosts or
pathogenic potential. This is particularly true under recent findings
that the diversification of bat coronaviruses is driven largely by host
shifts (inter-genus or higher levels of cross-species transmission) and,
to a lesser degree, cospeciation and sharing, representing intra-genus
cross-species transmission.\textsuperscript{2} This diversification is
not an actual risk factor for emergence itself, but acts downstream of
an emergence event by increasing the random chance of the emergence of a
virus with the raw genomic components required to eventually infect
humans.

\begin{figure}
\hypertarget{fig:risk}{%
\centering
\includegraphics{figures/risk_map.png}
\caption{Extraction of a measure of \emph{Betacoronavirus} emergence
risk from bat hosts based on the colorimetric space from
fig.~\ref{fig:trivariate}. The risk is a composite measure of the color
value and angular distance to the yellow hue, as defined in the methods,
ranged in the unit space. Based on these analyses, South-Eastern Asia,
Madagascar, the Middle-East, and Africa below the Sahara desert have the
highest relative risk of emergence.}\label{fig:risk}
}
\end{figure}

@@

The relationship between the underlying pathogen pool and emergence risk
is mediated by both human-wildlife interfaces (the probability of
spillover) and opportunities for onward transmission (the probability
that spillovers become epidemics)\textsuperscript{1}. As a proxy for
both, we finally overlaid the risk component from the composite map (see
above) with the proportion of built land, as a proxy for a mix of
habitat disturbance, potential for bat synanthropy or contact with
bridge hosts like livestock,\textsuperscript{25,26} and human population
density and connectivity\textsuperscript{1,27,28}
(fig.~\ref{fig:compound}). Accounting for these factors, most of South
America and Europe are at comparatively lower risk, as--although densely
populated--settlements tend to be in areas with lower potential risk.
Conversely, regions like Malaysia and the northern coast of Australia
have a high evolutionary risk component, but should represent a
relatively lower effective risk due to low human density. However,
southeast Asia, the Indian subcontinent, and scattered hotspots in
sub-Saharan Africa are at high risk due to the overlap between human
populations and natural opportunities for cross-species transmission of
betacoronaviruses.

@@

Reassuringly, these predictions correspond to the geographic origins of
the three bat-origin coronaviruses that have recently emerged in human
populations. While available information puts the spillover of
SARS-CoV-2 in a live animal market in Wuhan, China, the ultimate origin
of the virus is almost certainly in a divergent lineage of
sarbecoviruses from the Indochinese peninsula that was poorly
characterized prior to the pandemic.\textsuperscript{29--31} Similarly,
the SARS-CoV outbreak began in Guangdong province in 2002, reaching
humans through small carnivore bridge hosts, but was eventually traced
back to a set of likely progenitor viruses found in cave-dwelling
horseshoe bats in Yunnan province;\textsuperscript{32} nearby, antibody
evidence has indicated human exposure to SARS-like
viruses.\textsuperscript{33} MERS-CoV was originally detected in Saudi
Arabia, accompanied by a nearly identical virus sequenced from an
Egyptian tomb bat (\emph{Taphozous perforatus}),\textsuperscript{34} but
is widespread in camels in East Africa and the Middle East, and may have
reached its bridge host decades earlier than originally
supposed;\textsuperscript{35} as a result, the geography of the original
bat-to-camel transmission is still widely regarded as uncertain. All of
these are broadly consistent with the risk factors we identify. Notably,
India and west Africa are additional hotspots that have yet to
experience the emergence of a bat coronavirus into human populations,
but may still be at risk---particularly given known gaps in bat
surveillance,\textsuperscript{20} and a dense population in both regions
with global connectivity. In any of these regions, surveillance on viral
reservoirs can be paired with targeted monitoring of high-risk human
populations (i.e., those with regular wildlife
contact)\textsuperscript{36} for maximum impact.

@@

Bats are important reservoir hosts for different classes of
microorganisms, many of which a threat to human
health.\textsuperscript{37,38} Chiropterans emerged around 64 million
years ago and are one of the most diverse mammalian orders, with an
estimated richness of more than 1400 species.\textsuperscript{39,40}
They exhibit a broad variety of habitat use, behaviour, and feeding
strategies, putting them at key positions in the delivery and
provisioning of several ecosystem services, tied to important
ecosystem-derived benefits to human.\textsuperscript{41} For example,
bats are an essential component of many seed-dispersal
networks.\textsuperscript{42} Over two-thirds of bats are know to be
either obligate or facultative insectivores, therefore actively
contributing for agricultural pest control,\textsuperscript{43,44} and
vectors of pathogens that put a risk on human
health.\textsuperscript{45,46} Because bats are globally distributed and
have a long evolutionary history, phylogeographic and biogeographic
approaches are required to shed light on the contemporary distribution
of coevolutionary processes between bats and the pathogens they host.
Not all areas in which bats, viruses, and human are co-occuring are
facing a risk of spillover towards human populations, and the areas in
which this risk exist may not be facing risks of the same nature and
magnitude.

Here, we propose a simple freamework with broad explanatory power that
helps contextualize discoveries like highly divergent nobecoviruses in
Madagascar and the previously-neglected adaptive radiation of
sarbecoviruses outside of southern China and throughout southeast Asia.
In doing so, it advances ecological theory beyond the current state of
the art for global maps of emergence risk. For example, previous studies
that have used host richness as proxy have predicted a high diversity of
unsampled bat viruses,\textsuperscript{18} bat
coronaviruses,\textsuperscript{2} and even specifically
betacoronaviruses\textsuperscript{19} in both the Amazon and southeast
Asia. While we find that both regions are characterized by highly
divergent host and viral communities, our framework identifies key
differences between the regions. We find that Latin America is a hotspot
of both host and viral distinctiveness, suggesting that this branch of
the bat-betacoronavirus complex may be undergoing independent
evolutionary dynamics from the rest of the global pool, but with limited
potential for viral diversification--- a finding that is supported by
previous work indicating a higher rate of codivergence in Latin
America.\textsuperscript{2} In contrast, in southeast Asia, host
richness and viral distinctiveness are high but sharing is low; this
suggests a different type of evolutionary dynamics that could generate
high local diversity of viruses through host switching and viral
recombination (see \emph{e.g.},\textsuperscript{13} as well as the
discovery of recombinant viruses that share genetic material from both
the SARS-CoV and SARS-CoV-2 branches of the Sarbecovirus
lineage).\textsuperscript{47} Both of these regions are priority areas
for sampling, especially given predictions that they contain many bat
hosts of undiscovered betacoronaviruses.\textsuperscript{19,20} However,
both the evolutionary and ecological aspects of emergence risk are
likely higher in southeast Asia---a fact that will only become more
relevant, as bats track shifting climates and exchange viruses with
other species, creating a hotspot of cross-species transmission unique
to the region.\textsuperscript{48}

The diversity and diversification potential of bats responds to
anthropogenic factors others than shifting climates.\textsuperscript{49}
Land use changes could significantly decrease bat suitability, notably
through effects on diet and availability of
habitats.\textsuperscript{50} As our results establish that the
diversification of bats betacoronaviruses happens on top of processes
affecting hosts, biogeographic variation in human population density and
anthropogenic disturbances may feed into co-evolutionary dynamics.
Increase in humans-hosts contacts also increase the risk of emergence of
novel diseases,\textsuperscript{51} so does the changes in landscape
connectivity at local/regional scales.\textsuperscript{52} This
represents a challenge for both conservation strategies and disease
ecology: some areas can a high emergence risk and more potential for the
acquisition of zoonotic viruses through bat-human
encounters.\textsuperscript{53} In particular, the challenge ahead lies
in the need to quantify actual exposure (and risk) accounting for
several transmission scenarios, including both direct and indirect bat -
human interactions, and feeding back into the provision of ecosystem
services by bats.

@@

We downloaded the data on bats hosts of \emph{Betacoronavirus} from
\texttt{https://www.viralemergence.org/betacov} on
Apr.~2022,\textsuperscript{19} and filtered it to ``known'' hosts
(established before the emergence of SARS-CoV-2) and ``novel'' hosts
(confirmed through sampling and competence assays since the initial data
collection). The original database was assembled by a combination of
data mining and literature surveys, including automated alerts on the
``bats'' and ``coronavirus'' keywords to identify novel empirical
evidence of bats-betacoronaviruses associations; this yielded a total of
126 known hosts, 47 of which were novel hosts.

@@

We downloaded the rangemap of every current bat species that was
classified as an empirically documented host of \emph{Betacoronavirus}
from the previous step, according to recent IUCN
data.\textsuperscript{54} The range maps were subsequently rasterized
using the \texttt{rasterize} function from
\texttt{GDAL}\textsuperscript{55} at a resolution of approximately
100kmx100km. For every pixel in the resulting raster where at least one
bat host of \emph{Betacoronavirus} was present, we extract the species
pool (list of all known bat hosts), which was used to calculate the
following risk assessment components: bat phylogenetic diversity, bat
compositional uniqueness, and predicted viral sharing risk.

@@

For every pixel, we measured Faith's Phylogenetic
Diversity\textsuperscript{56} based on a recent synthetic tree with
robust time calibration, covering about 6000 mammalian
species.\textsuperscript{57} Faith's PD measures the sum of unique
branches from an arbitrary root to a set of tips, and comparatively
larger values indicate a more phylogenetic diverse species pool. We
measured phylogenetic diversity starting from the root of the entire
tree (and not from Chiroptera); this bears no consequences on the
resulting values, since all branches leading up to Chiroptera are only
counted one per species pool, and (as we explain when describing the
assembly of the composite risk map), all individual risk components are
ranged in {[}0,1{]}. This measure incorporates a richness component,
which we chose not to correct for; the interpretation of the
phylogenetic diversity is therefore a weighted species richness, that
accounts for phylogenetic over/under-dispersal in some places.
